% !TEX root = 00_dmlr_compel.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%  Compel: Curating High Quality Pretraining Data for Language Models
%  via Compression Ratios
%  DMLR (Journal of Data-centric Machine Learning Research) submission
%
%  Rebuilt from the double-blind review PDF (compel.pdf, 2025-12-01) because no
%  local .tex source exists on this machine; de-anonymized per DMLR single-blind
%  policy. Author list from Brando's CV (cv_long.tex, "Preprint 2026" entry).
%  Style:   dmlr2e.sty  (official, from https://github.com/JmlrOrg/dmlr-style-file)
%  Bib:     compel_refs.bib (rebuilt from the PDF reference list; placeholder
%           entries "arXiv:2401.xxxxx"-style repaired with real metadata)
%  New for DMLR: Section 4 (theoretical grounding: Shannon / Kolmogorov / MDL),
%  contributions list, Broader Impact, fixed Fig. 3 caption/panel mismatch,
%  deduplicated examples figure (was Figs. 2 and 5), dropped the leftover ICLR
%  template placeholder appendix.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[twoside,11pt]{article}
\usepackage{dmlr2e}

\usepackage{amsmath,amsfonts,amssymb}
\usepackage{graphicx}
\usepackage{url}
\usepackage{booktabs}
\usepackage{xcolor}
\usepackage{enumitem}
\usepackage{CJKutf8} % the high-CR example document in Fig.~\ref{fig:examples} is bilingual (Chinese/English)

\graphicspath{{./}{./figures/}}

% Green = improvement over the unfiltered baseline, red = regression (matches the review PDF's tables).
\newcommand{\imp}[1]{\textcolor[HTML]{1A7F37}{\textbf{#1}}}
\newcommand{\reg}[1]{\textcolor[HTML]{B91C1C}{\textbf{#1}}}

% Required by dmlr2e.sty — fill in after the OpenReview submission is created
\def\openreview{\url{https://openreview.net/forum?id=XXXX}}

% Journal heading metadata — fill in at camera-ready
% \dmlrheading{Volume}{Year}{pages}{submitted}{published}{id}{authors}

\ShortHeadings{Curating Pretraining Data via Compression Ratios}{Obbad, Miranda, Hall, Schaeffer, Koyejo \& Liang}

\firstpageno{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\title{Curating High Quality Pretraining Data for Language Models via Compression Ratios}

% TODO(internal-review): confirm author order, middle-initial style, and email handles with Elyas before submitting.
\author{%
  \name Elyas Obbad$^{1}$ \AND
  \name Brando Miranda$^{1}$ \AND
  \name David L.\ W.\ Hall$^{1}$ \AND
  \name Rylan Schaeffer$^{1}$ \AND
  \name Sanmi Koyejo$^{1}$ \AND
  \name Percy Liang$^{1}$ \\
  \addr $^{1}$Department of Computer Science, Stanford University, Stanford, CA 94305, USA \\
  \texttt{\{eobbad, brando9, dlwh, rschaef, sanmi, pliang\}@cs.stanford.edu}
}

\editor{Handling Editor: TBD}

\maketitle

%=========================  Abstract  ==============================%
\begin{abstract}
% CS197 move 1 — problem motivation: data quality gates capability, curation at web scale is compute-prohibitive.
The quality of pretraining data determines the capabilities of language models, yet identifying high-quality data among billions of web documents remains computationally prohibitive.
% CS197 move 2+3 — prior assumption (quality signals need models/heuristics) and the flip: a model-free scalar suffices.
We introduce \textsc{Compel}, a simple and scalable data processing step that isolates high-quality text using lightweight, compression-based signals.
Our key insight is that the compression ratio of text serves as a robust, model-free proxy for information density: low compression ratios typically reflect repetitive or boilerplate content, whereas high ratios may indicate noisy or unnatural text (e.g., HTML spam or phone numbers).
% CS197 move 4 — instantiation: retain documents inside a calibrated compression-ratio band.
\textsc{Compel} improves dataset quality by retaining only those documents whose compression ratios fall within a chosen range, determined empirically from high-quality reference datasets, without relying on additional model training or heuristic classifiers.
% CS197 move 5 — evaluation with concrete numbers.
\textsc{Compel} improves benchmark performance by 0.4--1.1 points across leading open web-scale datasets---DCLM, FineWeb, and FineWeb-EDU---while requiring only a fraction of the computational resources of traditional filtering methods.
% CS197 move 6 — implication: a practical, compute-efficient complement to existing quality controls.
These results show that compression-based filtering is a practical, compute-efficient complement to prevailing quality controls, capable of boosting pretraining data quality at trillion-token scale.
\end{abstract}

\begin{keywords}
pretraining data curation, data quality, data filtering, compression ratio, LZ4, information density, minimum description length, large language models, data-centric machine learning
\end{keywords}

\input{01_introduction}
\input{02_related_work}
\input{03_cr_signal}
\input{04_theory}
\input{05_method}
\input{06_experiments}
\input{07_results}
\input{08_qualitative}
\input{09_discussion}
\input{10_conclusion}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Broader Impact Statement (required by DMLR)
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\impact{%
\textsc{Compel} filters pretraining corpora with a single interpretable scalar, which lowers the compute, energy, and financial cost of data curation and makes quality filtering auditable: every discarded document has a stated, reproducible reason---its compression ratio.
Two risks deserve attention.
First, the band used here is calibrated on predominantly English web corpora; applying the same thresholds to other languages, scripts, or code could disproportionately remove valid non-English or mixed-language content (the high-ratio example in Figure~\ref{fig:examples} is bilingual), so practitioners should recalibrate the band per language or domain and audit what is removed.
Second, any automated quality filter shapes the distribution of voices in the resulting model; although \textsc{Compel} avoids reference-corpus classifiers that can encode cultural preferences, its band is still derived from curated corpora and can inherit their choices.
All corpora used in this work are publicly available.
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Bibliography  (style plainnat is set by dmlr2e.sty — do NOT re-issue
% \bibliographystyle or bibtex errors with a duplicate \bibstyle)
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliography{compel_refs}

\end{document}
