% !TEX root = 00_dmlr_compel.tex
\section{Experiments}\label{sec:experiments}

\subsection{Pretraining Corpora}\label{sec:corpora}

We evaluate \textsc{Compel} on three widely-used corpora for language model pretraining:
\begin{itemize}[leftmargin=1.35em,topsep=2pt,itemsep=1pt]
\item \textbf{FineWeb} \citep{penedo2024fineweb}: A 15-trillion-token web crawl filtered with domain heuristics and safety constraints.
\item \textbf{FineWeb-EDU} \citep{penedo2024fineweb}: A 1.3-trillion-token educational subset extracted using a synthetic LLM classifier trained on LLaMA 3 70B-Instruct annotations.
\item \textbf{DCLM} \citep{li2024datacomp}: A 4-trillion-token corpus derived from Common Crawl, curated to support high-accuracy model training with limited compute.
\end{itemize}

For each corpus, we construct a \emph{\textsc{Compel}-filtered} variant by removing documents whose LZ4 compression ratio falls outside the calibrated range:
\begin{equation*}
\tau_{\min} = 0.65, \qquad \tau_{\max} = 0.80.
\end{equation*}
Compression is applied to raw byte-level representations using a single pass over the dataset.
Filtering is performed prior to tokenization and shuffling.

\subsection{Model Configurations}\label{sec:models}

For each \textsc{Compel}-filtered variant, we train two decoder-only Transformer models based on the LLaMA architecture \citep{touvron2023llama} with rotary position embeddings and grouped-query attention.
The \textbf{1.4B model} has 16 layers, a hidden size of 2048, intermediate size of 7168, 16 attention heads, and 8 key/value heads.
The \textbf{8B model} consists of 32 layers, a hidden size of 4096, intermediate size of 14336, 32 attention heads, and 8 key/value heads.
Both models are configured with a maximum sequence length of 4096.
These values were selected based on standard configurations shown to perform well at this scale in prior open-source models.
The 8B model follows the LLaMA 3 hyperparameters \citep{touvron2023llama}, which are tuned for stability and efficiency at high model capacity.
All models are trained from scratch without weight reuse or initialization from pretrained checkpoints.

\subsection{Training Protocol}\label{sec:training}

The \textbf{1.4B model} is trained for 10{,}000 steps using a batch size of 1024 and sequence length of 4096, totaling 42 billion tokens.
The \textbf{8B model} is trained for 40{,}000 steps under the same sequence length and batch size, totaling 167 billion tokens.
Training is performed on TPU v4-128 pods.

We use AdamW optimization for all models.
The 1.4B model uses a learning rate of $3 \times 10^{-4}$ with weight decay of 0.1.
% TODO(internal-review): 2e-3 for the 8B looks high vs LLaMA-3's ~3e-4; this value is transcribed
% from the Dec-2025 review PDF — confirm the actual run config with Elyas before camera-ready.
The 8B model follows the LLaMA 3 hyperparameters: a learning rate of $2 \times 10^{-3}$, weight decay 0.05, and 5000 warmup steps.
All models are trained with cosine decay schedules and dropout rate of 0.1.

\subsection{Evaluation Suite}\label{sec:evaluation}

We evaluate each model using the standardized \texttt{lm-eval-harness} framework \citep{biderman2024lessons}, following zero-shot protocols on a diverse set of 13 natural language processing benchmarks:
\begin{itemize}[leftmargin=1.35em,topsep=2pt,itemsep=1pt]
\item \textbf{Commonsense and Reasoning:} COPA \citep{roemmele2011choice}, CommonsenseQA \citep{talmor2018commonsenseqa}, PIQA \citep{bisk2020piqa}, ARC (Easy/Challenge) \citep{clark2018think}, AGIEval LSAT-AR \citep{zhong2023agieval}
\item \textbf{Reading Comprehension:} BoolQ \citep{clark2019boolq}, OpenBookQA \citep{mihaylov2018suit}, WinoGrande \citep{sakaguchi2021winogrande}, WSC273 \citep{levesque2012winograd}, LAMBADA \citep{paperno2016lambada}
\item \textbf{Multichoice QA:} HellaSwag (0-shot, 10-shot) \citep{zellers2019hellaswag}
\end{itemize}

\paragraph{Metrics.}
We report task-standard metrics for each dataset, primarily multiple-choice question-answering accuracy.
We aggregate results using unweighted macro-average accuracy over all tasks.
