% !TEX root = 00_dmlr_compel.tex
\section{Results}\label{sec:results}

We evaluate \textsc{Compel} across three widely-used, large-scale pretraining datasets---FineWeb, FineWeb-EDU, and DCLM---using two LLaMA-style model scales: 1.4 billion (1.4B) and 8 billion (8B) parameters.
We measure model performance using macro-average accuracy across a suite of 13 diverse downstream benchmarks (see Section~\ref{sec:evaluation}).
The detailed benchmark results are summarized in Table~\ref{tab:results-1p4b}, with key findings discussed below.

\subsection{Impact of Compression-Based Filtering at 1.4B Scale}\label{sec:results-1p4b}

\begin{table}[t]
\centering
\caption{\textbf{Results for 1.4B-parameter models across datasets.}
\textsc{Compel} improves on 9 out of 13 benchmarks for FineWeb, 9 out of 13 for FineWeb-EDU, and 11 out of 13 for DCLM.
Green marks improvements over the corresponding unfiltered baseline; red marks regressions.}
\label{tab:results-1p4b}
\small
\setlength{\tabcolsep}{4.5pt}
\begin{tabular}{lcccccc}
\toprule
 & \multicolumn{2}{c}{FineWeb} & \multicolumn{2}{c}{FineWeb-EDU} & \multicolumn{2}{c}{DCLM} \\
\cmidrule(lr){2-3}\cmidrule(lr){4-5}\cmidrule(lr){6-7}
Benchmark & Base & +\textsc{Compel} & Base & +\textsc{Compel} & Base & +\textsc{Compel} \\
\midrule
WSC273             & 0.571 & \imp{0.604} & 0.571 & \reg{0.550} & 0.630 & \imp{0.637} \\
Winogrande         & 0.528 & \imp{0.534} & 0.532 & \reg{0.525} & 0.528 & \imp{0.539} \\
PIQA               & 0.708 & \imp{0.712} & 0.687 & \imp{0.702} & 0.705 & \imp{0.710} \\
OpenBookQA         & 0.196 & \reg{0.192} & 0.244 & \imp{0.258} & 0.222 & \imp{0.230} \\
LAMBADA            & 0.369 & \imp{0.379} & 0.337 & \imp{0.339} & 0.499 & \imp{0.501} \\
HellaSwag (10-shot)& 0.362 & \imp{0.365} & 0.365 & \reg{0.364} & 0.371 & \imp{0.378} \\
HellaSwag (0-shot) & 0.365 & \imp{0.369} & 0.378 & \imp{0.384} & 0.375 & \imp{0.379} \\
COPA               & 0.630 & \imp{0.730} & 0.690 & \imp{0.730} & 0.690 & \imp{0.710} \\
CommonsenseQA      & 0.201 & \reg{0.188} & 0.216 & \reg{0.197} & 0.199 & \imp{0.207} \\
BoolQ              & 0.591 & \imp{0.593} & 0.571 & \imp{0.606} & 0.507 & \reg{0.462} \\
ARC-Easy           & 0.537 & \reg{0.536} & 0.623 & \imp{0.634} & 0.609 & \imp{0.614} \\
ARC-Challenge      & 0.230 & \reg{0.218} & 0.282 & \imp{0.296} & 0.276 & \reg{0.265} \\
AGIEval LSAT-AR    & 0.230 & \imp{0.239} & 0.248 & \imp{0.257} & 0.217 & \imp{0.222} \\
\midrule
\textbf{Macro avg.}& 0.424 & \imp{0.435} & 0.442 & \imp{0.449} & 0.466 & \imp{0.470} \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{FineWeb.}
As shown in Table~\ref{tab:results-1p4b}, \textsc{Compel}-filtered models improve accuracy on 9 out of 13 tasks compared to the heavily-filtered FineWeb baseline, yielding a notable macro-average accuracy increase of 1.1 points (from 42.4\% to 43.5\%).
The largest improvements occur in commonsense reasoning benchmarks, notably COPA (+10\%) and WSC273 (+3.3\%), suggesting that removing documents outside the calibrated compression ratio range effectively eliminates subtle forms of noise, enhancing models' reasoning capabilities.

\paragraph{FineWeb-EDU.}
Even though FineWeb-EDU is already filtered via strong classifier-based heuristics, \textsc{Compel} further enhances model performance.
As Table~\ref{tab:results-1p4b} indicates, \textsc{Compel} filtering leads to improved accuracy on 9 of 13 tasks, resulting in a 0.7-point increase in macro-average accuracy (from 44.2\% to 44.9\%).
Notable gains are observed on COPA (+4\%) and BoolQ (+3.5\%).

\paragraph{DCLM.}
DCLM represents the current gold standard in dataset quality.
Nevertheless, Table~\ref{tab:results-1p4b} reveals that \textsc{Compel} still offers improvements on 11 of 13 tasks, yielding a 0.4-point increase in macro-average accuracy (from 46.6\% to 47.0\%).

\subsection{Effectiveness of \textsc{Compel} at Larger Model Scale (8B)}\label{sec:results-8b}

Our 8B experiments demonstrate that \textsc{Compel}'s effectiveness persists at scale: it improves both macro and micro average accuracy despite the increased capacity and learning dynamics of larger models.
This suggests that \textsc{Compel}'s benefits are not confined to small or medium-scale settings.
Instead, compression-based filtering enhances training efficiency even when models have greater capacity to absorb noise, indicating that the quality improvements it introduces are fundamental rather than compensatory.
As model sizes continue to grow, such lightweight filters can play an increasingly important role in optimizing compute budgets while preserving or improving performance.

Task-level results reveal notable improvements in tasks such as BoolQ (+5.5\%) and CommonsenseQA (+2.6\%), despite minor regressions on ARC-Challenge ($-$2.4\%) and WSC273 ($-$1.1\%).
These results highlight that \textsc{Compel} filtering scales effectively with increased model size, preserving or enhancing model capabilities and improving data efficiency without added computational complexity.

\begin{table}[t]
\centering
\caption{\textbf{Results on 8B models trained on FineWeb.}
\textsc{Compel} preserves or slightly improves downstream accuracy (macro average $0.570 \rightarrow 0.572$, micro average $0.587 \rightarrow 0.593$).}
\label{tab:results-8b}
\small
\begin{tabular}{lcc}
\toprule
Benchmark & FineWeb & FineWeb + \textsc{Compel} \\
\midrule
WSC273             & 0.809 & \reg{0.798} \\
Winogrande         & 0.682 & \reg{0.679} \\
PIQA               & 0.793 & \reg{0.791} \\
OpenBookQA         & 0.310 & 0.310 \\
LAMBADA            & 0.638 & \imp{0.648} \\
HellaSwag (10-shot)& 0.556 & \imp{0.561} \\
HellaSwag (0-shot) & 0.562 & 0.562 \\
COPA               & 0.860 & \reg{0.850} \\
CommonsenseQA      & 0.189 & \imp{0.215} \\
BoolQ              & 0.683 & \imp{0.738} \\
ARC-Easy           & 0.705 & \imp{0.732} \\
ARC-Challenge      & 0.396 & \reg{0.372} \\
AGIEval LSAT-AR    & 0.204 & \imp{0.213} \\
\midrule
\textbf{Macro avg.} & 0.570 & \imp{0.572} \\
\textbf{Micro avg.} & 0.587 & \imp{0.593} \\
\bottomrule
\end{tabular}
\end{table}
